\subsection{Part A: The AI recruitment dilemma}

\indent

You are a software engineer at TechHire Pty Ltd, building an AI system that screens job applicants. During testing, you discover:

\begin{itemize}
    \item The AI consistently ranks male candidates higher than female candidates, despite input data being designed to be gender-neutral.
    \item Management is pressuring you to launch the platform immediately, citing major client contracts and potential revenue losses if delayed.
    \item Public backlash could occur if discriminatory practices are exposed.
\end{itemize}

\bigskip

Each group in the tutorial has to pick one of the ethical models:

\noindent \textbf{Consequentialism (outcome-focused)}:

Action: Delay the launch until the bias is corrected.

Justification: The overall negative consequences of launching a biased system—legal issues, reputational damage, and harm to applicants—outweigh the short-term business benefits.

ACS Principles Applied:

\begin{itemize}
    \item Public Interest: Avoid harm to candidates and society.
    \item Honesty: Ensure the system functions as intended without discrimination.
    \item Quality of Life: Ensure fairness and equitable treatment.
\end{itemize}

\noindent\textbf{Deontology (rule/duty-focused)}:

Action: Do not launch the system as it violates ethical duties, such as fairness and non-discrimination.

Justification: Certain actions (like releasing a biased system) are inherently wrong, regardless of outcomes. Duties and rules, legal and professional, must be followed.

ACS Principles Applied:

\begin{itemize}
    \item Honesty: Do not misrepresent the system’s capabilities.
    \item Professional Competence: Ensure work meets ethical and technical standards.
    \item Public Interest: Respect rights and dignity of applicants.
\end{itemize}

\noindent\textbf{Virtue Ethics (character-focused)}

Action: Advocate for correcting the AI bias before launch, demonstrating integrity, fairness, and responsibility.

Justification: A virtuous IT professional acts according to moral character, showing honesty, courage, and fairness in decision-making.

ACS Principles Applied:
\begin{itemize}
    \item Honesty and Integrity: Act with moral character and transparency.
    \item Professional Competence: Deliver reliable and unbiased systems.
    \item Public Interest: Ensure the technology benefits society ethically.
\end{itemize}

Each group needs to discuss:

\begin{itemize}
    \item What action should be taken?
    \item How does your ethical model justify this decision?
    \item Which ACS principles guide your decision?
\end{itemize}

\bigskip

As a tutorial, have a group discussion:

\noindent\textbf{How do the decisions differ between models?}

Consequentialism emphasizes overall outcomes, deontology emphasizes following ethical duties, and virtue ethics emphasizes acting according to moral character. While all recommend delaying launch, the reasoning differs: outcomes vs. duty vs. 
character.

\noindent\textbf{Are there conflicts between professional ethics and business pressure?}

Yes, management’s pressure to launch conflicts with professional duties to ensure fairness, avoid harm, and uphold public trust. Resolving this conflict requires prioritizing ethical and professional standards over short-term business gains.

\subsection{Part B: Identifying the Ethical Issues}

For the following topics, identify the ethical issues and list out the relevant ACS principles:

\begin{itemize}
    \item AI in Healthcare: An AI system predicts patient risk but misclassifies some high-risk patients.
    \item Big Data in Marketing: Personalized ads use sensitive data without users’ explicit consent.
    \item IoT Smart Homes: Devices record audio/video constantly, sometimes without awareness.
    \item Cloud Data Storage: Sensitive client data is stored on servers in a different country with weaker privacy laws.
    \item Deepfakes: A celebrity’s likeness is used in an ad without permission.
\end{itemize}

\begin{table}[htbp]
\centering
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{|p{3cm}|p{6cm}|p{6cm}|}
\hline
\textbf{Technology} & \textbf{Ethical Issue} & \textbf{Relevant ACS Principles} \\
\hline

AI in Healthcare &
Misclassification of high-risk patients can lead to harm or delayed care. &
Public Interest, Professional Competence, Honesty, Quality of Life \\
\hline

Big Data in Marketing &
Using sensitive personal data without explicit consent violates privacy. &
Public Interest, Honesty, Privacy and Confidentiality, Professional Competence \\
\hline

IoT Smart Homes and Devices &
Continuous recording may invade privacy and confidentiality. &
Privacy and Confidentiality, Public Interest, Honesty, Professional Competence \\
\hline

Cloud Data Storage &
Storing sensitive data overseas may breach local laws and reduce control over data security. &
Privacy and Confidentiality, Public Interest, Professional Competence, Legal Compliance \\
\hline

Deepfakes &
Using a celebrity's likeness without consent may cause reputational harm and legal issues. &
Honesty, Public Interest, Professional Competence, Intellectual Property Respect \\
\hline

\end{tabular}
\caption{Ethical Issues and Relevant ACS Principles}
\end{table}