\subsection{PART A – Sources of Information}

\indent 

Go through the documentations under the section “Finding, Evaluating, and Referencing sources of information” on Canvas:

\begin{itemize}
    \item Watch the video on the page, evaluating sources of information.
    \item Read the following pages, finding sources of information, referencing sources of information, writing paragraphs and referencing sources in the text
\end{itemize}

\subsection{PART B – Concept Review}

\indent 

Discuss the following concepts in pairs or small groups:

\noindent\textbf{Knowledge}: 


Theoretical or practical understanding of a subject (e.g., understanding how databases work)

\noindent\textbf{Skill}: 

The ability to perform tasks well, often gained through practice (e.g., writing SQL queries)

\noindent\textbf{Expertise}: 

Deep, refined knowledge and skill in a specific area, usually with experience (e.g., being a data architect with 10+ years of experience)

\noindent\textbf{Profession}: 

A field of work requiring formal education, training, and adherence to ethical standards (e.g., law, medicine, IT).

\noindent\textbf{Professionalism}: 

The conduct and qualities that characterize a professional, such as accountability, integrity, and competence.

\noindent\textbf{Practice}: 

The application of knowledge and skills in real-world scenarios over time (e.g., consulting or managing IT systems).

\subsection{PART C – Being an IT Professional}

\noindent\textbf{What does it mean to be an IT professional in today’s world?}: 

\begin{itemize}
    \item Demonstrating responsibility, ethical behavior, and technical competence in the design, implementation, and management of IT systems. 
    \item Continuously learning and adapting to emerging technologies and business needs.
    \item Working collaboratively with others while upholding the values and standards of the profession.
\end{itemize}

\noindent\textbf{Identify three technical and three non-technical skills that are essential for IT professionals.}: 

\begin{itemize}
    \item Technical skills: Programming, Data analysis, Cybersecurity, Networking, etc.
    \item Non-technical skills: Communication, Critical thinking, problem solving, teamwork, collaboration, etc.
\end{itemize}


\noindent\textbf{What are some challenges that IT professionals face when trying to maintain professionalism?}: 

\begin{itemize}
    \item Balancing business needs with ethical decisions
    \item Navigating changing technologies and maintaining competence.
    \item Communicating effectively with non-technical stakeholders
\end{itemize}

\subsection{PART D – Role of IT in Organisations}

\noindent\textbf{What are some strategic roles that IT plays in modern organisations?}: 

\begin{itemize}
    \item Streamlining business operations through automation
    \item Enabling data-driven decision-making
    \item Supporting innovation and product/service delivery
    \item Enhancing customer experience and engagement
    \item Ensuring compliance and security
\end{itemize}

\noindent\textbf{Identify the four key forces reshaping how organisations operate and compete.}: 

\begin{itemize}
    \item Business drivers - Business goals like cost reduction, efficiency, scalability, and innovation push IT to deliver faster, cheaper, and smarter solutions.
    \item Technology drivers - Emerging technologies (e.g., AI, IoT, cloud, automation) shape how businesses operate and offer services, requiring IT professionals to adapt rapidly.
    \item Customer expectations - Demand for personalised, seamless, 24/7 digital experiences forces IT to focus on UX design, mobile access, security, and performance.
    \item Market Trends - Competitive pressure, new business models (like SaaS), and global shifts (e.g., remote work) require IT systems to be agile and market-responsive.
\end{itemize}

\noindent\textbf{For each force, give one real-world example of how it affects IT practice or business operations.}