\subsection{Assessments}

\noindent\textbf{Early Semester Feedback Task}

\begin{itemize}
    \item Multiple choice, True/False
    \item On Canvas – Open book
    \item Syllabus – Week 1 to 3
    \item Held in Week 4
\end{itemize}

\bigskip

\noindent\textbf{Viva}

\begin{itemize}
    \item Oral quiz held in tutorials
    \item Secured assessment – No notes allowed
    \item Syllabus – Week 1 to Week 7
    \item Held in Week 8
\end{itemize}

\bigskip

\noindent\textbf{Group Assignment}

\begin{itemize}
    \item Groups of 4-5 people
    \item "Topic of assessment"
    \item Submission in Week 12
\end{itemize}

\bigskip

\noindent\textbf{Final Exam}

\begin{itemize}
    \item Restrictive open book – 1 page both side
    \item Case Study based Questions
    \item Held in Exam week
\end{itemize}

\subsection{Basic Concept}

\subsubsection{Knowledge, Skills and Expertise}

\noindent\textbf{Knowledge}: Knowing about something, how to do it.

\noindent\textbf{Skills}: Being able to do it – competence, capability

\noindent\textbf{Expertise = Knowledge + Skills}

\subsubsection{Profession}

\noindent\textbf{Profession}: An occupation requiring specialized education, expertise, and practical skills. It is guided by ethical standards, often regulated by professional bodies, and serves the public by addressing complex societal needs.

\subsubsection{Practice}

\noindent\textbf{Practice}: refers to the repeated application of knowledge, skills, or activities to improve performance or maintain proficiency. It can also mean the actual work done within a profession or discipline.

\subsubsection{Professionalism}

\noindent\textbf{Professionalism}: is the conduct, behavior, and attitude expected of someone in a professional role. It reflects a commitment to competence, ethics, integrity, and respect in the workplace.

\bigskip

\noindent\textbf{Key aspects}:

\begin{itemize}
    \item Competence - having the necessary knowledge and skills.
    \item Accountability - taking responsibility for your actions.
    \item Ethical Behavior - acting with honesty, fairness, and integrity.
    \item Reliability - being dependable and consistent.
    \item Respect and courtesy - treating others with dignity and professionalism
\end{itemize}

\subsubsection{Being an IT professional}

\noindent \textbf{Being an IT professional} goes beyond just working with technology. It is about expertise, responsibility, and ethics in using technology to solve problems and create value.

\noindent\textbf{Core skills in today's world}: Tech Skills = Capability. Soft Skills = Adaptability. Tech Can Be Taught, Soft Skills Take Time.

\subsection{IT Careers}

\indent 

In the fast-evolving world of Information Technology, career paths can vary greatly depending on your experience, skills, and specialization. Whether you're a student just entering the field or a seasoned professional, job roles and responsibilities grow with you.

The next few slides will showcase real-world examples of IT job descriptions across different stages of an IT career, from entry-level roles for recent graduates to mid-level positions, and then senior or specialized roles.

This will help you understand how expectations change over time and what you can work toward as you grow in the industry

\subsubsection{Progression of IT careers}

\noindent\textbf{Year 0-2}: Entry level / Graduate roles

\begin{itemize}
    \item \textbf{Job Titles}: IT Support, Junior Developer, Data Analyst, QA Tester, Business Analyst
    \item \textbf{Focus}: Building foundational technical skills, learning industry tools, working under supervision
    \item \textbf{Skills Gained}: Coding, troubleshooting, system usage, documentation, teamwork
\end{itemize}

\bigskip

\noindent\textbf{Year 3-5}: Intermediate / Associate roles

\begin{itemize}
    \item \textbf{Job Titles}: Software Engineer, Systems Analyst, DevOps Engineer, Data Engineer, Cybersecurity Analyst
    \item \textbf{Focus}: Managing projects/components independently, starting to mentor juniors, deeper domain knowledge
    \item \textbf{Skills Gained}: Full-stack development, cloud computing, process automation, security practices
\end{itemize}

\bigskip

\noindent\textbf{Year 6-8}:

\begin{itemize}
    \item \textbf{Job Titles}: Senior Developer, Solutions Architect, Tech Lead, Product Manager, Senior Analyst
    \item \textbf{Focus}: Leading projects or teams, shaping systems or product direction, engaging stakeholders
    \item \textbf{Skills Gained}: Leadership, system design, cross-functional collaboration, technical decision-making
\end{itemize}

\bigskip

\noindent\textbf{Year 9-12+}:

\begin{itemize}
    \item \textbf{Job Titles}: IT Manager, Engineering Manager, Enterprise Architect, CTO, CIO
    \item \textbf{Focus}: Organizational strategy, team management, innovation, aligning IT with business goals
    \item \textbf{Skills Gained}: Strategic planning, budgeting, stakeholder management, organizational leadership
\end{itemize}

\bigskip

\noindent\textbf{Alternate Paths}:

\begin{itemize}
    \item Consulting - Moving into IT consulting or advisory roles
    \item Research - Pursuing a PhD or research-intensive career in AI, HCI, etc.
    \item Entrepreneurship - Founding or joining a startup
    \item Specialization - Becoming an expert in niche areas (e.g., cybersecurity, ML/AI, blockchain)
\end{itemize}

\subsubsection{Typical activities in IT careers}

\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|c|cl|}
\hline
\textbf{Technical}     & \textbf{\begin{tabular}[c]{@{}c@{}}Analytical\&\\ Problem solving\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Project\&Process\\ Management\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Collaboration \&\\ Communication\end{tabular}} & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Learning \&\\ Innovation\end{tabular}}} \\ \hline
\textbf{Coding}        & Troubleshooting                                                                 & Project planning                                                               & Stakeholder comm                                                                  & \multicolumn{2}{c|}{Continuous learn}                                                          \\ \hline
\textbf{Testing}       & System design                                                                   & Agile/Scrum                                                                    & Team collaboration                                                                & \multicolumn{2}{c|}{Evaluating tools}                                                          \\ \hline
\textbf{Debugging}     & Data collection                                                                 & Time management                                                                & User training                                                                     & \multicolumn{2}{c|}{Adopting innovation}                                                       \\ \hline
\textbf{Prototyping}   & Data analysis                                                                   & Cost management                                                                & Managing vendors                                                                  & \multicolumn{2}{c|}{Knowledge share}                                                           \\ \hline
\textbf{Cybersecurity} & Reporting                                                                       & Quality assurance                                                              & Client interactions                                                               & \multicolumn{2}{c|}{Innovating}                                                                \\ \hline
\end{tabular}
\end{table}

\subsection{IT in Organizations}

\subsubsection{Role of IT}

\indent

The \textbf{role of IT} (Information Technology) is to enable, support, and transform the way individuals, businesses, and governments \textbf{store, process, communicate, and access information}.

It serves as the backbone of modern operations and innovation.

\bigskip

\noindent\textbf{Core roles of IT}:

\begin{itemize}
    \item Enabling Communication and collaboration
    \item Data management and storage
    \item Process automation
    \item System development and maintenance
    \item Cybersecurity and risk management
    \item Supporting innovation and digital transformation
    \item Technical support and troubleshooting
\end{itemize}

\subsubsection{Role of IT in organizations}

\indent

IT provides value to an organization through changing the way in which business functions and processes are carried out.

IT provides value by providing new functions and enabling new business models.

\subsection{Changing Business Landscape}

\noindent The 4 key forces that reshape how organizations operate and compete:

\begin{itemize}
    \item Business drivers
    \item Technology drivers
    \item Customer expectations
    \item Market trends
\end{itemize}

\subsection{Question}

\subsubsection{End of lecture questions}

\begin{itemize}
    \item Define the following terms and explain how they differ: Profession, Practice, Professionalism
    \item What is the purpose of INFO5990, and how does it prepare students for IT careers?
    \item List three ways technological advancement has changed the nature of IT careers.
    \item Briefly explain the role of IT in modern organizations. Give two examples.
    \item Describe three soft skills that are essential in IT careers and explain why they are important.
\end{itemize}

\subsubsection{Case-study based Questions}

\indent

Alex is a recent IT graduate who just started a job as a junior systems analyst at a large consulting firm. In their first few weeks, they realized that although they had strong technical skills, they struggled with stakeholder communication and time management. Their team lead recommended that they revisit the concepts taught in INFO5990 to improve their workplace professionalism and soft skills.

\bigskip

\noindent\textbf{Question}:

\begin{itemize}
    \item Identify two aspects of professionalism that Alex is currently struggling with. Justify your answer.
    \item How could Alex’s understanding of “professional practice” help him improve his performance and confidence at work?
    \item Suggest three strategies Alex can implement to enhance his soft skills, drawing from INFO5990 learning.
    \item Explain how changing business landscapes demand new skillsets in IT roles, and how Alex can prepare for that.
\end{itemize}